### 2014-10-29 stat374/hw2.tex
started homework 2

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%% AMS PACKAGES - Chances are you will want some or all of these if writing a math dissertation.
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, enumerate, multicol, graphicx, listings}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Homework 2 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[a)]
                I will use SURE to estimate the risk of the generalized James-Stein estimator. Let
                \[\hat\theta^{JS,v}= Z - \frac{(Z-v)(n-2)\sigma^2}{\sum_{i=1}^n (Z_i - v)^2}= Z + g(Z) \]  
                Accordingly, 
                \begin{align*}
                    D_i &= \frac{\partial g_i}{\partial z_i} \\
                    D_i  &= -(n-2)\sigma^2\left(\frac{1}{\sum_{i=1}^n (z_i - v)^2} - \frac{(z_i-v)^2}{\left(\sum_{i=1}^n 2(z_i - v)^2\right)^2}\right) \\
                    \sum_{i=1}^n D_i &= -(n-2)\sigma^2\left(\frac{n}{\sum_{i=1}^n (z_i - v)^2} - \frac{2\sum_{i=1}^n(z_i-v)^2}{\left(\sum_{i=1}^n (z_i - v)^2\right)^2}\right) \\
                    \sum_{i=1}^n D_i &= -(n-2)\sigma^2\left(\frac{n-2}{\sum_{i=1}^n (z_i - v)^2}\right)  \\
                    \sum_{i=1}^n D_i &= -\frac{(n-2)^2\sigma^2}{\sum_{i=1}^n (z_i - v)^2} 
                \end{align*}
                Plugging this into the SURE formula, we get that
                \begin{align*}
                    \hat R (Z) &= n\sigma^2 + 2\sigma^2 \sum_{i=1}^n D_i + \sum_{i=1}^n g_i^2 \\
                    \hat R (Z) &= n\sigma^2 - 2\frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} + \sum_{i=1}^n \left(\frac{-(Z_i-v)(n-2)\sigma^2}{\sum_{i=1}^n (Z_i - v)^2}\right)^2 \\
                    \hat R (Z) &= n\sigma^2 - 2 \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} + \frac{(n-2)^2\sigma^4\sum_{i=1}^n (Z_i - v)^2}{\left(\sum_{i=1}^n (Z_i - v)^2\right)^2} \\
                    \hat R (Z) &= n\sigma^2 - 2 \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} + \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} \\
                    \hat R (Z) &= n\sigma^2 - \frac{(n-2)^2\sigma^4}{\sum_{i=1}^n (Z_i - v)^2} \\
                \end{align*}
                Since the estimate is unbiased,
                \[ R(\hat\theta^{JS,v},\theta) = \E(\hat R(Z)) = n\sigma^2 - (n-2)^2\sigma^4\E\left(\frac{1}{\sum_{i=1}^n (Z_i - v)^2}\right) \]
                Since \(Z_i-v \sim N(\theta_i - v,\sigma^2)\), we can use a result from the book that 
                \[\E\left(\frac{1}{\sum_{i=1}^n (Z_i - v)^2}\right) \geq \frac{1}{(n-2)\sigma^2 + \|\theta-v\|^2} \]
                This gives us that
                \begin{align*}
                    R(\hat\theta^{JS,v},\theta) &\leq n\sigma^2 - \frac{(n-2)^2\sigma^4}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq \frac{n\sigma^2((n-2)\sigma^2 + \|\theta-v\|^2) - (n-2)^2\sigma^4}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq \frac{n(n-2)\sigma^4 + n\sigma^2\|\theta-v\|^2 - (n-2)^2\sigma^4}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq \frac{2\sigma^2(n-2)\sigma^2 + (2 + n -2)\sigma^2\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                    R(\hat\theta^{JS,v},\theta) &\leq 2\sigma^2 + \frac{(n -2)\sigma^2\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2} \\
                \end{align*}
                Because 
                \[1 < \frac{\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2}\]
                We get that, for $n\geq3$,
                \[R(\hat\theta^{JS,v},\theta) \leq 2\sigma^2 + \frac{(n -2)\sigma^2\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2} < 2\sigma^2 + (n-2)\sigma^2 = n\sigma^2 \]
                \(R(\hat\theta^{JS,v},\theta)\) will be smallest when \(\frac{\|\theta-v\|^2}{(n-2)\sigma^2 + \|\theta-v\|^2}\) is smallest, which in turn is when \(\|\theta-v\|^2\) is smallest. Thus, $v=\bar\theta$, the mean of $\theta$, will result in the smallest risk.
            \item[b)]
                In the case where $n=1$, since $\hat\theta = Z$ is admissiable, there is no estimator $\tilde\theta$ such that \(R(\tilde\theta,\theta)<R(\hat\theta,\theta)\), so $\hat\theta^{a,b}$ is admissible only if 
                \[R(\hat\theta,\theta)=\E(\hat\theta-\theta)^2=\E(\hat\theta^{a,b}-\theta)^2=R(\hat\theta^{a,b},\theta)\]
                This implies that, to be admissiable, it must be the case that $Z=aZ+b$ for all potential $Z$. Thus, since this is not the case for i, ii, iii, and iv, we must conclude the estimator is not admissable in these cases. Only in case iv can $Z=aZ+b$ for all potential $Z$, so there the estimator is admissable. 
                \smallskip
                As we showed in part a, $R(Z,\theta)=n\sigma^2>R(\hat\theta^{JS,v},\theta)$ when $n\geq3, so $\hat\theta = Z$ is not admissable. However, there are certainly cases where \(R(\hat\theta^{JS,v},\theta)\) is not admissable. Take for instance the example of $\theta_i=c$ for some constant $c$. The mean of $Z$ of has risk \(R(\bar Z,\theta) = \frac{\sigma^2}{n} < \)

        \end{itemize}
\end{itemize}
\end{document}