### 2015-04-07 stat302/hw1.tex
Started homework 1

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.5in]{geometry}
\pdfpagewidth 8.5in
\pdfpageheight 11in
\setlength\parindent{0in}

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
STAT 302 Homework 1 \hfill Aaron Maurer
\vspace{2mm}
\hrule
\vspace{2mm}

\begin{itemize}
    \item[1.]
        \begin{itemize}
            \item[(a)]
                We can calculate the type I error as
                \[P(x\in\{1,2\}|\theta=0) = P(x=1|\theta=0) + P(x=2|\theta=0) = .01\]
                And the type II error as 
                \[P(x\not\in\{1,2\}|\theta=1) = P(x=3|\theta=1) =.01\]
            \item[(b)]
                We can calculate the p-value for $x=1$ as 
                \[ \sum_{x: P(X=x|H_0)\leq P(X=1|H_0)} \hspace{-4em} P(X=x|H_0) = P(X=1|H_0) + P(X=2|H_0) = .01\]
                The Bayes Factor for $H_1$ vs. $H_2$ is
                \[\frac{P(x=1|\theta=1)}{P(x=1|\theta=0)} =\frac{.0049}{.005} = .98\]
                Qualitatively, this means that $x=1$ is very, very slightly evidence for $H_0$ over $H_1$.
            \item[(c)]
                We get a p-value for $x=2$ of 
                \[ \sum_{x: P(X=x|H_0)\leq P(X=2|H_0)} \hspace{-4em} P(X=x|H_0) = P(X=1|H_0) + P(X=2|H_0) = .01\]
                But this time the Bayes Factor $H_1$ vs. $H_2$ is
                \[\frac{P(x=2|\theta=1)}{P(x=2|\theta=0)} =\frac{.9851}{.005} = 197.02\]
                Which is overwhelming evidence for $H_1$
            \item[(d)]
                If we modify our example to \\
                \begin{table}[h!]
                \center
                \begin{tabular}{c|c|c|c}
                $x$ & 1 & 2 & 3 \\ \hline
                $p(x|\theta=0)$ & 0.005 & 0.005 & 0.99 \\
                $p(x|\theta=1)$ & 0.495 & 0.495 & 0.01 
                \end{tabular}
                \end{table}

                Then the p-value for $x=1$ is still .01, but now the Bayes Factor is
                \[\frac{P(x=1|\theta=1)}{P(x=1|\theta=0)} =\frac{.495}{.005} = 99\]
                Which is a very large and in favor of $H_1$
        \end{itemize}
    \item[2.]

    \item[3.]
        To get the posterior, we do the usual Bayesian calculation:
        \begin{align*}
            P(\mu|x) &=  P(\mu) P(x|\mu) \\
                     &\propto \mu^{m-1}e^{-\lambda \mu} \prod_{i=1}^n \frac{\mu^{x_i}}{x_i!}e^{-\mu} \\
                     &\propto \mu^{m-1+\sum_{i=1}^n x_i}e^{-(\lambda+n) \mu} \prod_{i=1}^n \frac{1}{x_i!} \\
        \end{align*}
\end{itemize}

\end{document}