### 2015-06-08 stat302/paper_report.tex
Started on paper report; finished overview

\documentclass[11pt]{article}
\usepackage[paper=letterpaper, margin=.75in]{geometry}
\linespread{1.5}
\pdfpagewidth 8.5in
\pdfpageheight 11in

%%% Packages
% First four - AMS (american mathematical society). General math goodness. I use the align* enviorment in particular
% multirow, multicol allow for certain kinds of tables
% enumerate lets you determine the style of the counter for the enumerate enviorment
% graphicx lets you include pictures
% listings lets you stick in blocks of code
% placeins defines "\FloatBarrier", which stops tables from moving around
\usepackage{amsmath, amscd, amssymb, amsthm, multirow, multicol, enumerate, graphicx, listings, placeins} 
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\U}{\mathcal{U}}
\newcommand{\del}{\partial}
\newcommand{\real}{\textrm{Re }}
\newcommand{\imag}{\textrm{Im }}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\deriv}[2]{\frac{d #1}{d #2}}
\newcommand{\sumk}{\sum_{k=1}^\infty}
\newcommand{\sumj}{\sum_{j=1}^\infty}
\newcommand{\sumn}{\sum_{n=0}^\infty}
\newcommand{\summ}[2]{\sum_{k=#1}^{#2}}
\newcommand{\sig}[1]{\sum_{#1 =1}^\infty}
\newcommand{\un}[1]{\bigcup_{#1 =1}^\infty}
\newcommand{\inter}[1]{\bigcap_{#1 =1}^\infty}
\newcommand{\ip}[2]{\langle #1, #2 \rangle}
\newcommand{\ipxu}{\langle x,u_j \rangle}
\newcommand{\uj}{\{u_j\}_{j=1}^\infty}
\newcommand{\B}{\mathcal{B}}

\newcommand{\E}{\mathrm{E}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\ST}{mbox{ s.t. }}

\newcommand{\Example}{\noindent {\bf Example. \quad} }
\newcommand{\Proof}{\noindent {\bf Proof: \quad} }
\newcommand{\Remark}{\noindent {\bf Remark. \quad} }
\newcommand{\Remarks}{\noindent {\bf Remarks. \quad} }
\newcommand{\Case}{\noindent {\underline{Case} \quad} }

\newcommand{\st}{ \; \big | \:}

\newcommand{\deuc}{d_{\mathrm euc}}
\newcommand{\dtaxi}{d_{\mathrm taxi}}
\newcommand{\ddisc}{d_{\mathrm disc}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\newcommand{\hwhead}[1]{#1 \hfill Aaron Maurer \vspace{2mm} \hrule \vspace{2mm}}

\begin{document}
\title{Literature Report on DP-means}
\author{Aaron Maurer}
\date{STAT 302, Spring Quarter 2015}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
For this report, I chose to read "Revisiting k-means: New Algorithms via Bayesian Nonparametrics"\footnote{Brian Kulis and Michael I. Jordan, ``Revisiting k-means: New Algorithms via Bayesian Nonparametrics'', \textit{CoRR} (2011): http://arxiv.org/abs/1111.0352}. This article introduces a alternative to the k-means clustering algorithm called DP-means, where instead of specifying a number of clusters at the start, then working to optimize their fit, new clusters are introduced when a point in the data set is too far from one of the existing clusters. This in itself is a `hard' clustering algorithm, which outputs an assignment to clusters, rather than a posterior distribution one might normally expect from a Bayesian method. However, this paper demonstrates that this algorithm is the limit of a Bayesian model where the total distribution is a mixture of Gaussian distributions, with the number of components coming from a Dirichlet process. This is the source of the name of the algorithm: Dirichlet process means or DP-means. \par
In addition to a description and derivation of this algorithm, the paper offers a few extensions and simulations. The most interesting of theses is a hierarchical clustering method where, over multiple data sets of the same variables, local clusters are simultaneously fit on each data set so as to match a set of global clusters across all of the data sets. This is again a hard clustering algorithm, but it can be derived as well in a Bayesian fashion by taking the limit when the parameters of the Dirichlet process which generates the components in each data set arises from a prior global Dirichlet process. The next extension shows how DP-means extends to spectral clustering; where one would spectrally cluster using k-means by performing k-means on the first k eigenvectors of a similarity matrix for the data, instead with DP-means one takes all eigenvectors where the eigenvalues are above a given threshold, then cluster via DP-means using that same threshold to determine when to add additional components. Similarly, the author also shows that DP-means can be extended to graph cut problems. The paper concludes with a few simulations demonstrating the effectiveness of DP-means and its multiple data set extension. \par
This article offers obvious extensions of how the classical Bayesian statistics we learned can be extended to a machine learning algorithm. The prior Dirichlet process is used to guide the posterior number of components. The method for fitting the model is the limit of a Gibbs sampling algorithm. Finally, the multiple data set version of DP-means is a hierarchical model, based on the exchangeability of the data sets. All together, this forms an interesting competitor to k-means, being similarly easy to compute, but built on classic Bayesian principles.

\section{Paper Contents}
\subsection{DP-Means Algorithm}

\end{document}